---
title: "Per Protocol Analysis"
author: "Sam Jaros"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup}
library(tidyverse)

pp_data = readRDS(here::here("data/primary_data.rds")) %>%
  mutate(pod1_med_log = if_else(pod1_med==0,0,log(pod1_med)),
         fu_med_log = if_else(fu_med==0,NA_real_,log(fu_med))) %>%
  filter(tx_group=="Control" | fu_tx_used)
```

## Per Protocol Analysis

To see if the treatment effects are even more extreme for those who reported using the treatment, I have restricted the treated group to only those who reported using the treatment at 1 month. The control group is the same.

```{r restricted_takingOp}
takingOp_pp_basic = glm(fu_taking_opioids~tx_group + pod1_med_log,
                        family="binomial", data=pp_data)
# summary(takingOp_pp_basic)
exp(coefficients(takingOp_pp_basic)[2])
exp(confint(takingOp_pp_basic, 2))

takingOp_pp_demog = glm(fu_taking_opioids~tx_group + age + raceeth + gender +
                          pod1_med_log,
                        family="binomial", data=pp_data)
# summary(takingOp_pp_demog)
exp(coefficients(takingOp_pp_demog)[2])
exp(confint(takingOp_pp_demog, 2))

takingOp_pp_full = glm(fu_taking_opioids~tx_group + age + raceeth + gender +
                          pod1_med_log + n_surg_types + length_of_stay,
                        family="binomial", data=pp_data)
# summary(takingOp_pp_full)
exp(coefficients(takingOp_pp_full)[2])
exp(confint(takingOp_pp_full, 2))
```

All of the confidence intervals cross 1 suggesting there is no significant effect of the treatment on odds of taking opioids at all when controlling for using the treatment at 1 month.

## Part 2: Amount of Opioids Taken

```{r restricted_amtOp}
amtOp_pp_basic = glm(fu_med_log ~ tx_group + pod1_med_log,
                     family=gaussian, data=pp_data)
# summary(amtOp_pp_basic)
coef(amtOp_pp_basic)[2]
confint(amtOp_pp_basic, 2)

amtOp_pp_demog = glm(fu_med_log ~ tx_group + age + raceeth + gender +
                       pod1_med_log,
                     family=gaussian, data=pp_data)
# summary(amtOp_pp_demog)
coef(amtOp_pp_demog)[2]
confint(amtOp_pp_demog, 2)

amtOp_pp_full = glm(fu_med_log ~ tx_group + age + raceeth + gender +
                          pod1_med_log + n_surg_types + length_of_stay,
                     family=gaussian, data=pp_data)
# summary(amtOp_pp_full)
coef(amtOp_pp_full)[2]
confint(amtOp_pp_full, 2)
```

All of the confidence intervals cross zero which suggests there is no statistically significant effect of the treatment on amount of opioids taken when controlling for using the treatment at 1 month.
